---
title: "p8105_mtp_rw2844"
author: "Renjie Wei"
date: "10/23/2020"
output: 
  github_document:
    toc: TRUE
---

```{r setup, include = FALSE}
library(tidyverse)
library(readxl)
library(patchwork)
library(ggridges)
library(rvest)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 10, 
  fig.height = 8,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis", 
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Report

## Problem 1

First, I imported the data.
```{r read_data}
dog_weight = 
  read_excel("./data/dogweights_07sept2020.xlsx", 
    sheet = "Sheet1", range = "B2:G207", 
    col_types = c("date", "text", "text", 
        "text", "text", "text")) %>% 
  janitor::clean_names() 
colnames(dog_weight) = c("date", "raisin", "simone", "gagne", "std", "notes")
# a missing value in date
```

In this chunk, I cleaned and manipulated the data, selected columns except the notes, the major steps are:

* Separate the weight data into different units(Lbs and Oz).

* Do some calculation, generate a new variable to combine Lbs and Oz into one unit.

* Make the dataset looks better using `pivot_longer`.

* Select the variables of interest.
```{r frame_1}
dogs_df = 
  dog_weight %>%
  # Noticing that weight of 3 dogs are in two different unit, we need to split it into two columns
  separate(
    col = c("raisin"),
    into = c("raisin_lbs", "raisin_oz"),
    sep = (2)
  ) %>% 
  separate(
    col = c("simone"),
    into = c("simone_lbs", "simone_oz"),
    sep = (2)
  ) %>% 
  separate(
    col = c("gagne"),
    into = c("gagne_lbs", "gagne_oz"),
    sep = (2)
  ) %>% 
  separate(
    col = c("std"),
    into = c("std_lbs", "std_oz"),
    sep = (2)
  ) %>% 
  # change variables type
  mutate(
    raisin_lbs = as.numeric(raisin_lbs),
    raisin_oz = as.numeric(raisin_oz),
    simone_lbs = as.numeric(simone_lbs),
    simone_oz = as.numeric(simone_oz),
    gagne_lbs = as.numeric(gagne_lbs),
    gagne_oz = as.numeric(gagne_oz),
    std_lbs = as.numeric(std_lbs),
    std_oz = as.numeric(std_oz)
  ) %>% 
  # assuming that 1 lbs equals 16 oz, combine two rows
  # noticing that there is a outlier in raisin's weight(7 Lbs 15 Oz), it might be a typo, I'm gonna to fix it
  mutate(
    raisin = raisin_lbs + (raisin_oz/16),
    simone = simone_lbs + (simone_oz/16), 
    gagne = gagne_lbs + (gagne_oz/16),
    std = std_lbs + (std_oz/16),
    raisin = case_when(
      raisin >= 10 ~ raisin,
      raisin < 10 ~ raisin+10
    )
  ) %>% 
  # make the data looks better
  pivot_longer(
    raisin:std,
    names_to = "object",
    values_to = "weight_lbs"
  ) %>% 
  mutate(date = as.Date(date), object = as.factor(object)) %>% 
  # remain what we need
  select(date, object, weight_lbs)

# check for variable types
str(dogs_df)

# show the dataframe
dogs_df %>% 
  head(12) %>% 
  knitr::kable(digits = 2, caption = "The weight of three dogs(Lbs), first 12 rows")
```

In the next chunk, I am going to create a dataframe that only includes notes and dates.

```{r frame_2}
notes_df = 
  dog_weight %>% 
  select(date, notes) %>% 
  drop_na(notes)

notes_df %>% 
  knitr::kable(caption = "Notes and dates")
```

Save these two dataframes.

```{r save_data}
write.csv(dogs_df,"./data/dogs_df.csv")
write.csv(notes_df,"./data/notes_df.csv")
```