---
title: "p8105_mtp_rw2844"
author: "Renjie Wei"
date: "10/23/2020"
output: 
  github_document:
    toc: TRUE
---

```{r setup, include = FALSE}
library(tidyverse)
library(readxl)
library(patchwork)
library(ggridges)
library(rvest)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 10, 
  fig.height = 8,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis", 
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Report

## Problem 1

First, I imported the data.
```{r read_data}
dog_weight = 
  read_excel("./data/dogweights_07sept2020.xlsx", 
    sheet = "Sheet1", range = "B2:G207", 
    col_types = c("date", "text", "text", 
        "text", "text", "text")) %>% 
  janitor::clean_names() 
colnames(dog_weight) = c("date", "raisin", "simone", "gagne", "std", "notes")
# a missing value in date
```

In this chunk, I cleaned and manipulated the data, selected columns except the notes, the major steps are:

* Separate the weight data into different units(Lbs and Oz).

* Do some calculation, generate a new variable to combine Lbs and Oz into one unit.

* Make the dataset looks better using `pivot_longer`.

* Select the variables of interest.
```{r frame_1}
dogs_df = 
  dog_weight %>%
  # Noticing that weight of 3 dogs are in two different unit, we need to split it into two columns
  separate(
    col = c("raisin"),
    into = c("raisin_lbs", "raisin_oz"),
    sep = (2)
  ) %>% 
  separate(
    col = c("simone"),
    into = c("simone_lbs", "simone_oz"),
    sep = (2)
  ) %>% 
  separate(
    col = c("gagne"),
    into = c("gagne_lbs", "gagne_oz"),
    sep = (2)
  ) %>% 
  separate(
    col = c("std"),
    into = c("std_lbs", "std_oz"),
    sep = (2)
  ) %>% 
  # change variables type
  mutate(
    raisin_lbs = as.numeric(raisin_lbs),
    raisin_oz = as.numeric(raisin_oz),
    simone_lbs = as.numeric(simone_lbs),
    simone_oz = as.numeric(simone_oz),
    gagne_lbs = as.numeric(gagne_lbs),
    gagne_oz = as.numeric(gagne_oz),
    std_lbs = as.numeric(std_lbs),
    std_oz = as.numeric(std_oz)
  ) %>% 
  # assuming that 1 lbs equals 16 oz, combine two rows
  # noticing that there is a outlier in raisin's weight(7 Lbs 15 Oz), it might be a typo, I'm gonna to fix it
  mutate(
    raisin = raisin_lbs + (raisin_oz/16),
    simone = simone_lbs + (simone_oz/16), 
    gagne = gagne_lbs + (gagne_oz/16),
    std = std_lbs + (std_oz/16),
    raisin = case_when(
      raisin >= 10 ~ raisin,
      raisin < 10 ~ raisin+10
    )
  ) %>% 
  # make the data looks better
  pivot_longer(
    raisin:std,
    names_to = "object",
    values_to = "weight_lbs"
  ) %>% 
  mutate(date = as.Date(date), object = as.factor(object)) %>% 
  # remain what we need
  select(date, object, weight_lbs)

# check for variable types
str(dogs_df)

# show the dataframe
dogs_df %>% 
  head(12) %>% 
  knitr::kable(digits = 2, caption = "The weight of three dogs(Lbs), first 12 rows")
```

In the next chunk, I am going to create a dataframe that only includes notes and dates.

```{r frame_2}
notes_df = 
  dog_weight %>% 
  select(date, notes) %>% 
  drop_na(notes)

notes_df %>% 
  knitr::kable(caption = "Notes and dates")
```

Save these two dataframes.

```{r save_data}
write.csv(dogs_df,"./data/dogs_df.csv")
write.csv(notes_df,"./data/notes_df.csv")

```


# Problem 2

For the tidy dataset `dogs_df`, there are `r nrow(dogs_df)` rows and `r ncol(dogs_df)` variables, the number of unique dates is `r n_distinct(pull(dogs_df, date))`, including `r sum(is.na(unique(pull(dogs_df,date)))) ` missing value in `date`.



```{r cleaned_data}
dogs_df %>% 
  filter(object != "std") %>%
  # noticing that there are NA in the weight_lbs, so we cannot use summarize directly
  count(object, weight_lbs, name = "counts") %>% 
  drop_na(weight_lbs) %>% 
  group_by(object) %>% 
  summarise(
    n_obs = sum(counts),
    mean_weight = mean(weight_lbs),
    sd_weight = sd(weight_lbs)
  ) %>% 
  knitr::kable(digits = 2, caption = "Summarize for the three dogs weight")
```

# Problem 3

In the next chunk, I create a two-panel plot showing:

* In the left panel, the distribution of weights for each dog.

* In the right panel, each dog’s weight over time.
```{r plot}

plot_a = 
  dogs_df %>% 
    filter(object != "std") %>%
    # noticing that there are NA in the weight_lbs, so we cannot use summarize directly
    count(object, weight_lbs, name = "counts") %>% 
    drop_na(weight_lbs) %>% 
  ggplot(aes(y = weight_lbs, fill = object)) +
  geom_histogram(alpha = 0.5, binwidth = 0.3) +
  geom_density(alpha = 0.2) +
  labs(
    title = "Distribution of Weights for Each Dog",
    x = "Count / Frequncey",
    y = "Weight(Lbs)",
    fill = "Object" 
  )
 
  
  

plot_b =
  dogs_df %>% 
    filter(object != "std") %>%
  ggplot(aes(x = date, y = weight_lbs, color = object))+
  geom_point(alpha = 0.3) +
  geom_smooth(se = F) +
  labs(
    title = "Dog’s Weight Over Time",
    x = "Date(Y/M/D)",
    y = "Weight(Lbs)",
    color = "Object" 
  )


plot_a + plot_b
```
  
The plots above show the distribution of weight and the weight change for the dogs. From the left panel, we can see that the distribution of weight for each dog is pretty like a bell shape, which means the distribution of their weights are approximately normal. From the right panel, we can see the trend of the change in their weight. Obviously, there is a significant decrease in Gagne's weight. That is reasonable because Gagne is the only one who died among all three dogs, he may had a bad health condition. The weight of the other two dogs changes smoothly and steadily over time, indicating that their physical condition is relatively good. In addition, there is also a slight decrease in Raisin's weight during the first few months.
